---
title: "Activity 8"
author: "Name"
output: github_document
---

## Data and packages

Load the entirety of the `{tidyverse}`.
You will do a little work with dates in this activity so also load `{lubridate}`.
Be sure to avoid printing out any unnecessary information and give the code chunk a meaningful name.

```{r}
library(tidyverse)
```


Cheatsheets that you might want to add to your collection:

- [`{stringr}`](https://stringr.tidyverse.org/index.html)
- [`{forcats}`](https://forcats.tidyverse.org/)

Using `here::here`, upload the `billboard_songs.txt` file that is saved in your `data/` folder.
Notice that this file is a tab-delimited file that is stored as a `.txt` file.
Therefore, you will need to use either `read_delim` with a `delim = ...` argument or (better) `read_tsv`.
Assign the file to a meaningful object name, be sure to avoid printing any unnecessary information, and give the code chunk a meaningful name.

```{r}
setwd("~/STA518/Preps/activity08-special-variables")
df = read_tsv(here::here('data','billboard_songs.txt'))
```



These data include information on song popularity.
In the US, the Billboard Hot 100 is a list that comes out every week, showing the 100 most played songs that week.
More information about the creation of this dataset, as well as some analyses by the author, can be found here: [https://mikekling.com/analyzing-the-billboard-hot-100/](https://mikekling.com/analyzing-the-billboard-hot-100/).
The dataset you are provided is a limited version of the full data, containing:

- `title`
- `artist`
- `overall peak`: The highest rank the song ever reached (1 is the best)
- `weeks on chart`: The number of weeks the song was on the chart
- `chart date`: The latest date the song appeared on the Billboard Hot 100

This is a long dataset (34,605 observations)!

## Analysis

### Cleaning date

When you look at ``chart date`` you'll notice that it was read in as a numeric (`<dbl>`) variable type.
What does the format of the date appear to be in?
```{r}
head(df$`chart date`,10)
```

**Response**:
yyyymmdd, and may need to change to character if using stringr

In this activity, we will be using `str_...` functions to split or find patterns in strings of information.
Do the following with your dataset:

- *Create* a variable called `date` that treats ``chart date`` *as a character* variable type,
- Use `stringr::str_sub` on `date` to create three new columns: `year`, `month`, `day`
- Replace/overwrite the variable `date` with an updated format using `lubridate::make_date`.

Once you have verified that this worked, overwrite your data object.
```{r}
df$date = as.character(df$`chart date`)
df$y = str_sub(df$date,1,4)
df$m = str_sub(df$date,5,6)
df$d = str_sub(df$date,7,8)
```

```{r}
df$date2 = lubridate::as_date(df$date)
```



#### Analyzing using dates

What 10 songs spent the longest on the charts?
Give only the title, artists, and weeks.

```{r}
df$weeks=df$`weeks on chart`
df%>%
  select('title','artist','weeks')%>%
  slice_max(order_by = weeks,n=10)
  
```



**Response**:
```{r}
library(lubridate)
```


What date did the oldest song(s) in this dataset leave the charts?
Give only the *distinct* dates.

```{r}
df$date3=as.numeric(df$date2)
df$date4=df$date3+7*df$weeks
df$dateOut = lubridate::as_date(df$date4)
```

```{r}
df%>%
  select('title','artist','dateOut')%>%
  slice_min(order_by = df$date4,n=10)
```


**Response**:


### Artists who work together

Before we get to far, let's create common definitions:

In the string below, Nicki Minaj and Young Thug are considered to be **featured**.

```
RAE SREMMURD featuring NICKI MINAJ & YOUNG THUG
```

In the string below, Jessie J and Ariana Grande and Nicki Minaj are all considered to have **collaborated** on the song.

```
JESSIE J, ARIANA GRANDE & NICKI MINAJ
```

Which artist has been **featured** on the most Billboard charting songs?

First, create a vector object called `artist_credentials` that takes your dataset and *then* `pull`s only the `artist` column.
The `pull` function is like `select` except that it outputs a vector rather than a data frame (tibble).
Using this function will help us avoid many "`coercing`" warnings - I find it easier to work with vectors of strings.
```{r}
try=df %>%
  select(artist)%>%
  filter(str_detect(artist,'featuring'))
```
<<<<<<< HEAD

```{r}
try2=df %>%
  select(artist)%>%
  filter(str_detect(artist,'featuring'))%>%
  pull(artist)

```

```{r}
get_after_period <- function(my_vector) {    

        # Return a string vector without the characters
        # before a period (excluding the period)

        # my_vector, a string vector

        str_sub(my_vector, str_locate(my_vector, "\\.")[,1]+1) 

        }
```


```{r}
a=try2[1]
a
b=str_replace_all(a,"featuring",".")
get_after_period(b)
```


```{r}
b=str_replace_all(try2,"featuring",".")
try3=get_after_period(b)
try3[1:10]
```
=======

```{r}
try2=df %>%
  select(artist)%>%
  filter(str_detect(artist,'featuring'))%>%
  pull(artist)

```
Using the `artist_credentials` vector, do the following. I recommend that you do one bullet at time and verify that it worked.
>>>>>>> 05f51ba58ee3e0a0dc6cf0b25372f84c759f778b

- Use `str_split` to separate entries with multiple featured artist - these could include `,`, `&`, ` or ` (note the whitespace), or ` and ` (note the whitespace) (this should produce a list), *then*
- Use `unlist` to lengthen all featured artists into one vector (compare the output here to the output from the previous bullet point), *then*
- Use `str_trim` to remove whitespace from `"both"` sides of each entry.


```{r}
try4 = str_split(try3,'[,|&|or|and]')
try5 = str_trim(try4,side=c("both"))
```


```{r}
try6=str_count(try5)
```


```{r}
df1 = data.frame(try5,try6)
df1%>%
  select(try5,try6)%>%
  slice_max(order_by = try6,n=5)
```



